What we {\sf eval} in the shadows

### 2021-04-01 paper/main.tex
5

\section{A Taxonomy of Eval}

@@

\eval has three arguments, two of which we study in more details:

\begin{itemize}
	\item \c{expr}: the expression to evaluate
	\item \c{envir}: the environment in which performing the evaluation
	\item \c{enclos}
\end{itemize}

\subsubsection{The expression passed to \eval.}
\paragraph{Touching package environments.}
It refers to the \emph{namespace} of a package. A call to
\c{library(packageName)}, which load libraries in R, for instance, will populate
environment \c{package:packageName}. It is a \emph{base} function though, so we
do not see it in the CRAN dataset. However, R still makes it possible to access
any package environments and \packageNbPackageEnvSites call sites, \ie
\packagePackageEnvSitePercent access a package environment, from
\packageNbPackageEnvPackages different packages. Those packages perform
introspection, to represent and search any environment by name, in package
\emph{envnames}, to containerize tests, as with \emph{testthat} and
\emph{unitizer}, to create a custom OOP system with \emph{R.oo}, or to create
poor man's hidden functions in \emph{USSCXenaTools}.

It is also used to implement a custom caching mechanism in packages
\emph{g.data} and \emph{SOAR}.

During tracing, we gather all source expressions and all resolved expression
arguments passed to \eval. For instance, a non-resolved expression and some of
its resolved expressions can be:

\begin{lstlisting}
parse(text = paste0(\"options(\", i, \"=\", opts[i], \")\"))

options(datatable.alloccol = 1024L)
options(datatable.allow.cartesian = FALSE)
options(datatable.print.keys = FALSE)
\end{lstlisting}

@@

\begin{table}[h]\small
\begin{tabular}{|c|c|c|c|c|c|}\hline
Minimized & \#sites & \%sites & \#packages & \#operations & \%envir \\	\hline
\texttt{X} & \packageMinimizedcallsitesa  & \packageMinimizedpropsitesa & \packageMinimizedpackagea & \packageMinimizedoperationsaRnd & \packageMinimizedpercentenvirnonparenta \\\hline
\texttt{F(F())} & \packageMinimizedcallsitesb  & \packageMinimizedpropsitesb & \packageMinimizedpackageb  & \packageMinimizedoperationsbRnd & \packageMinimizedpercentenvirnonparentb \\\hline
\texttt{V} & \packageMinimizedcallsitesc & \packageMinimizedpropsitesc & \packageMinimizedpackagec & \packageMinimizedoperationscRnd & \packageMinimizedpercentenvirnonparentc \\\hline
\texttt{F(X)}& \packageMinimizedcallsitesd & \packageMinimizedpropsitesd & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & \packageMinimizedpercentenvirnonparentd \\\hline
\texttt{model.frame} & \packageMinimizedcallsitese & \packageMinimizedpropsitese & \packageMinimizedpackagee & \packageMinimizedoperationseRnd & \packageMinimizedpercentenvirnonparente \\\hline
\texttt{F()}& \packageMinimizedcallsitesf & \packageMinimizedpropsitesf & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & \packageMinimizedpercentenvirnonparentf \\\hline
\texttt{\$}& \packageMinimizedcallsitesg & \packageMinimizedpropsitesg & \packageMinimizedpackageg & \packageMinimizedoperationsgRnd & \packageMinimizedpercentenvirnonparentg \\\hline
\texttt{FUN} & \packageMinimizedcallsitesh & \packageMinimizedpropsitesh & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & \packageMinimizedpercentenvirnonparenth \\\hline
\texttt{<-} & \packageMinimizedcallsitesi  & \packageMinimizedpropsitesi & \packageMinimizedpackagei & \packageMinimizedoperationsiRnd & \packageMinimizedpercentenvirnonparenti \\\hline
\texttt{\{BLOCK\}} & \packageMinimizedcallsitesj & \packageMinimizedpropsitesj & \packageMinimizedpackagej & \packageMinimizedoperationsjRnd & \packageMinimizedpercentenvirnonparentj \\\hline
\texttt{F(F(<-))} & \packageMinimizedcallsitesk & \packageMinimizedpropsitesk & \packageMinimizedpackagek & \packageMinimizedoperationskRnd & \packageMinimizedpercentenvirnonparentk \\\hline
\end{tabular}
\caption{Minimized expressions: eleven most frequent in number of sites. \#sites
  is the number of sites; \%sites the percentage of sites across the total
  number of sites (\packageNbCallSites sites) with at least this minimized
  expression; \#packages is the number of packages with at least one of the
  minimized expression and measures how widespread a minimized expression is.
  \#operations is the average number of steps performed by the R interpreter
  when given that minimized expression, and \%envir is the percentage of call
  sites for which the \c{envir} argument is explicitly passed, and is not
  \c{parent.frame()}, which is the most common explicit environment passed to
  \eval.} \label{tab:minimizedexpressions}
\end{table}

@@

\paragraph{Base.} The ranking of minimized expressions in \emph{base} is similar to the one in CRAN bu the share of \c{V} is much larger, as it is present in \baseMinimizedpropsitesa\% of the sites and \c{X} is a distant second, in \baseMinimizedpropsitesb\% of the sites.

\paragraph{Kaggle.} Kaggle is dominated by sites with minimized expression \c{V}, with \kaggleMinimizedcallsitesa sites, and  \kaggleMinimizedpropsitesa\% of the \kaggleNbCallSites sites in total. All of those values are an AST size of one node, suggesting they do not come. \NOTE{TODO}
%TODO: all the src ref here seem to be incorrect and not from kaggle.

>>>>

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Taxonomy of Eval}

@@

%%%%%%%% JAN:::: NOT SURE WHAT THIS IS
%% It refers to the \emph{namespace} of a package. A call to
%% \c{library(packageName)}, which load libraries in R, for instance, will populate
%% environment \c{package:packageName}. It is a \emph{base} function though, so we
%% do not see it in the CRAN dataset. However, R still makes it possible to access
%% any package environments and \packageNbPackageEnvSites call sites, \ie
%% \packagePackageEnvSitePercent access a package environment, from
%% \packageNbPackageEnvPackages different packages. Those packages perform
%% introspection, to represent and search any environment by name, in package
%% \emph{envnames}, to containerize tests, as with \emph{testthat} and
%% \emph{unitizer}, to create a custom OOP system with \emph{R.oo}, or to create
%% poor man's hidden functions in \emph{USSCXenaTools}.

%% It is also used to implement a custom caching mechanism in packages
%% \emph{g.data} and \emph{SOAR}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The argument passed to a call to \eval determines what the function will do. As
we have seen in the previous section, there are cases where the argument is a
value, but these are rare. The common cases are that the argument is some form
of expression, either a single variable, a simple expression or a statement.

To categorize the expressions, consider a function $min(e)$ which for a given
expression $e$ returns a normal form that abstracts many incidental details
allowing the reader to focus on the structure.

@@

\begin{table}[h]\small
\begin{tabular}{|c|r|r|r|r|r|c|}\hline
$min(e)$& \#sites & \%sites & \#packages & \#operations & \%envir & example\\\hline
\c X&\packageMinimizedcallsitesa &\packageMinimizedpropsitesa &\packageMinimizedpackagea &\packageMinimizedoperationsaRnd &\packageMinimizedpercentenvirnonparenta & \c{y+1}\\\hline
\c{F(F())} & \packageMinimizedcallsitesb  & \packageMinimizedpropsitesb & \packageMinimizedpackageb  & \packageMinimizedoperationsbRnd & \packageMinimizedpercentenvirnonparentb & \c{gbov( mean(x), a-1)}\\\hline
\c{V}&\packageMinimizedcallsitesc &\packageMinimizedpropsitesc &\packageMinimizedpackagec &\packageMinimizedoperationscRnd &\packageMinimizedpercentenvirnonparentc&c(42,21,0)\\\hline
\c{F(X)}& \packageMinimizedcallsitesd & \packageMinimizedpropsitesd & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & \packageMinimizedpercentenvirnonparentd \\\hline
\c{model.frame} & \packageMinimizedcallsitese & \packageMinimizedpropsitese & \packageMinimizedpackagee & \packageMinimizedoperationseRnd & \packageMinimizedpercentenvirnonparente \\\hline
\c{F()}& \packageMinimizedcallsitesf & \packageMinimizedpropsitesf & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & \packageMinimizedpercentenvirnonparentf \\\hline
\c{\$}& \packageMinimizedcallsitesg & \packageMinimizedpropsitesg & \packageMinimizedpackageg & \packageMinimizedoperationsgRnd & \packageMinimizedpercentenvirnonparentg \\\hline
\c{FUN} & \packageMinimizedcallsitesh & \packageMinimizedpropsitesh & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & \packageMinimizedpercentenvirnonparenth \\\hline
\c{<-} & \packageMinimizedcallsitesi  & \packageMinimizedpropsitesi & \packageMinimizedpackagei & \packageMinimizedoperationsiRnd & \packageMinimizedpercentenvirnonparenti \\\hline
\c{\{\}} & \packageMinimizedcallsitesj & \packageMinimizedpropsitesj & \packageMinimizedpackagej & \packageMinimizedoperationsjRnd & \packageMinimizedpercentenvirnonparentj \\\hline
\c{F(F(<-))} & \packageMinimizedcallsitesk & \packageMinimizedpropsitesk & \packageMinimizedpackagek & \packageMinimizedoperationskRnd & \packageMinimizedpercentenvirnonparentk \\\hline
\end{tabular}
\caption{Minimized expressions: eleven most frequent in number of sites. \#sites
  is the number of sites; \%sites the percentage of sites across the total
  number of sites (\packageNbCallSites sites) with at least this minimized
  expression; \#packages is the number of packages with at least one of the
  minimized expression and measures how widespread a minimized expression is.
  \#operations is the average number of steps performed by the R interpreter
  when given that minimized expression, and \%envir is the percentage of call
  sites for which the \c{envir} argument is explicitly passed, and is not
  \c{parent.frame()}, which is the most common explicit environment passed to
  \eval.} \label{tab:minimizedexpressions}
\end{table}

@@

\subsubsection{The expression passed to \eval.}
\paragraph{Touching package environments.}

\paragraph{Base}
The ranking of minimized expressions in \emph{base} is similar to the one in
CRAN bu the share of \c{V} is much larger, as it is present in
\baseMinimizedpropsitesa\% of the sites and \c{X} is a distant second, in
\baseMinimizedpropsitesb\% of the sites.

\paragraph{Kaggle}
Kaggle is dominated by sites with minimized expression \c{V}, with
\kaggleMinimizedcallsitesa sites, and \kaggleMinimizedpropsitesa\% of the
\kaggleNbCallSites sites in total. All of those values are an AST size of one
node, suggesting they do not come. \NOTE{TODO}
%TODO: all the src ref here seem to be incorrect and not from kaggle.
